---
title: 'Working Out: You’re Doing it Wrong'
author: "HT"
date: "March 16, 2015"
output: html_document
        keep_md: true
---
```{r}
install.packages("dplyr", repos="http://cran.rstudio.com/")
library("dplyr")
install.packages("caret", repos="http://cran.rstudio.com/")
library("caret")
install.packages("caret", repos="http://cran.rstudio.com/")
library("caret")
```

#Executive Summary


#Get and Process the Data
The first step of the analysis to retrieve the data from the internet, save it to a local folder and load it into R.

```{r cache=TRUE}
download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="train_031215.csv", method="curl")
download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile="test_031215.csv", method="curl")

train <- read.csv("train_031215.csv")
test <- read.csv("test_031215.csv")
```

Then we divide the data into training and validation sets.
```{r}
inTrain <- createDataPartition(y=train$classe, p=0.9, list=FALSE)
rawTrain <- train[inTrain,]
vTrain <- train[-inTrain,]
```

We will need to pare away predictors with little value. To begin, we identify predictors with high numbers of NAs or no measurements at all. After this the dataset only has 53 predictors. To futher eliminate unnecessary predictors, random forest (from the caret package) was run to determine the most important variables.

```{r cache=TRUE}
x <- sapply(rawTrain, function(x) sum(is.na(x))) ##Sums the number of NAs in each row.
x <- x[x >= 19216] ## creates a vector of only those with 19,216 NAs.
x <-names(x) ## creates a vector with each column name
rawTrain <- rawTrain[,!(names(rawTrain) %in% x)] ## eliminates all columns that appear in rawTrain and in the object x
rawTrain <- select(rawTrain, -1, -c(3:7), -c(12:20), -c(43:48), -c(52:61), -c(74:82)) ## eliminates rows with lots of blanks using 'select' from the dplyr package.
trialRF <- train(classe ~ ., rawTrain, method="rf", importance=TRUE, prox=TRUE)
varImp(trialRF)
```

Of the top 20 variables listed none after the first 8 have an importance value > 50.0 for any of the 5 classes. So those variables are captured as the maximum number of potential predictors.

```{r}
fit1 <- glm(classe ~ roll_belt + pitch_belt + pitch_forearm + magnet_dumbbell_y + magnet_dumbbell_z + yaw_belt + accel_forearm_x + roll_forearm, data = rawTrain, family="quasibinomial")
summary(fit1)
```

A quick test, looking at the significance of the variables in a generalized linear model (family = "quasibinomial" i.e., "logit") reveals that one of them, "magnet_dumbell_y," has no significance in the presence of the other variables, so it to is cut from the prediction process (above). 

```{r}
finalTrain <- select(rawTrain, classe, roll_belt, pitch_belt, pitch_forearm, magnet_dumbbell_z, yaw_belt, accel_forearm_x, roll_forearm)
```

Random Forest was run again in order to create the best route given the new variable set.

```{r}
finalRF <- train(classe ~ ., finalTrain, method="rf", importance=TRUE, prox=TRUE)
```

Then we use the model to predict the values in the validation set:

```{r}
pred <- predict(finalRF, vTrain)
vTrain$right <- pred==vTrain$classe
table(pred, vTrain$classe)
```

We can see that the chosen model was able to accurately predict X values of the 1960 values in the training set, an accuracy rating of Y. Although not perfect, the accuracy is quite good and the model was adopted for predicting the values in the test set. Of the 20 cases in the test set, the chosen model predicted all 20 correctly.
 ============================================================
The following code was used to create text files for submitting predictions
of the test cases to Coursera for evaluation.
pml_write_files = function(x){
n = length(x)
for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

###Citation:
This dataset is licensed under the Creative Commons license (CC BY-SA). The CC BY-SA license means you can remix, tweak, and build upon this work even for commercial purposes, as long as you credit the authors of the original work and you license your new creations under the identical terms we are licensing to you. This license is often compared to "copyleft" free and open source software licenses. All new works based on this dataset will carry the same license, so any derivatives will also allow commercial use.

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
